\documentclass{sig-alternate-2013}

\newfont{\mycrnotice}{ptmr8t at 7pt}
\newfont{\myconfname}{ptmri8t at 7pt}
\let\crnotice\mycrnotice%
\let\confname\myconfname%

\permission{Permission to make digital or hard copies of all or part of
    this work for personal or classroom use is granted without fee
    provided that copies are not made or distributed for profit or
    commercial advantage and that copies bear this notice and the full
    citation on the first page. Copyrights for components of this work
    owned by others than the author(s) must be honored. Abstracting with
    credit is permitted. To copy otherwise, or republish, to post on
    servers or to redistribute to lists, requires prior specific
    permission and/or a fee. Request permissions from
permissions@acm.org.}
\conferenceinfo{CCS'13,}{November 4--8, 2013, Berlin, Germany. \\
    {\mycrnotice{Copyright is held by the owner/author(s). Publication
rights licensed to ACM.}}}
\copyrightetc{ACM \the\acmcopyr}
\crdata{978-1-4503-2477-9/13/11\ ...\$15.00.\\
    http://dx.doi.org/10.1145/2508859.2516697 
}

\clubpenalty=10000 
\widowpenalty = 10000

\usepackage{url}
\newcommand{\fixme}{{\bf XXX\ }}

\begin{document}

%don't want date printed
\date{}

\title{Tappan Zee (North) Bridge: Mining Memory Accesses for Introspection}

\numberofauthors{4}

\author{
\alignauthor
Brendan Dolan-Gavitt\\
    \affaddr{Georgia Tech}\\
    \email{brendan@cc.gatech.edu}
\alignauthor
Tim Leek\\
    \affaddr{MIT Lincoln Laboratory}\\
    \email{tleek@ll.mit.edu}
\alignauthor
Josh Hodosh\\
    \affaddr{MIT Lincoln Laboratory}\\
    \email{josh.hodosh@ll.mit.edu}
\and
\alignauthor
Wenke Lee\\
    \affaddr{Georgia Tech}\\
    \email{wenke@cc.gatech.edu}
} % end author

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
\thispagestyle{empty}

\begin{abstract}

The ability to introspect into the behavior of software at runtime is
crucial for many security-related tasks, such as virtual machine-based
intrusion detection and low-artifact malware analysis. Although some
progress has been made in this task by automatically creating programs
that can passively retrieve kernel-level information, two key challenges
remain. First, it is currently difficult to extract useful information
from user-level applications, such as web browsers. Second, discovering
points within the OS and applications to hook for \emph{active
monitoring} is still an entirely manual process. In this paper we
propose a set of techniques to mine the memory accesses made by an
operating system and its applications to locate useful places to deploy
active monitoring, which we call \emph{tap points}. We demonstrate the
efficacy of our techniques by finding tap points for four useful
introspection tasks such as finding SSL keys and monitoring web browser
activity on five different operating systems (Windows 7, Linux, FreeBSD,
Minix and Haiku) and two processor architectures (ARM and x86).\footnote{This
work is sponsored by the Assistant Secretary of Defense for Research \&
Engineering under Air Force Contract \#FA8721-05-C-0002. Opinions,
interpretations, conclusions and recommendations are those of the author and
are not necessarily endorsed by the United States Government.}

\end{abstract}

\category{D.4.6}{Operating Systems}{Security and Protection}
\terms{Design, Security}
\keywords{Introspection; active monitoring; security}

\input{introduction}

\input{tapdef}

\input{scope}

\input{technical}

\input{implementation}

\input{evaluation}

\input{limitations}

\input{relwork}

\input{conclusion}

\input{acks}

\bibliographystyle{abbrv}
\bibliography{biblio}

\input{appendix}

\end{document}
